{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-05 thesis/landslide.tex
explain the landslide approach

Landslide is a model checker implemented as a plug-in module for x86 full-system simulators.
The program to be tested runs in a simulated environment,
and Landslide uses its access to the simulator's internal state to inspect and manipulate the memory and thread scheduling of the program as it executes.
As of this thesis's writing, Landslide supports the use of two possible simulators:

@@

@@

\subsubsection{State machine}

The first part of {\tt sched\_update()} is to update the state machine of thread actions and runnability.
Much of this functionality is found in
{\tt sched\_update\_kern\_state\_machine()}
and
{\tt sched\_update\_user\_state\_machine()}.
The current intruction pointer is compared against the known locations of
the mutex API, system calls, runnable/descheduling {\tt tell\_\allowbreak{}landslide} annotations, and so on,
and locksets, action flags, and runqueue membership are updated accordingly.

\subsubsection{Interrupt injection}

@@

@@

% time travel, explore, arbiter
\subsection{Bug-finding output}

@@

\subsubsection{Infinite loop detection}

@@

\subsubsection{Yield-loop detection}

@@

\subsubsection{False-positive deadlock avoidance}

>>>>

Landslide is a model checker implemented as a plug-in module for x86 full-system simulators.
The program to be tested runs in a simulated environment,
and Landslide uses its access to the simulator's internal state to inspect and manipulate the memory and thread scheduling of the program as it executes.
\revision{\Cref{fig:landslide-architecture} visualizes Landslide in relation to its execution environment,
showing how it communicates with each of its surrounding and/or simulated programs.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.875\textwidth]{landslide-new.pdf}
	\end{center}
	\caption{Landslide's execution environment when testing 15-410 student projects.}
	\label{fig:landslide-architecture}
\end{figure}

\subsubsection{Design}

From an implementation point of view,
Landslide's main ``execution loop'' is simply the simulated CPU's own fetch-decode-execute loop:
each time it simulates an instruction or memory access,
it invokes Landslide through the simulator's module interface.
More abstractly speaking, the general sequence of events during a Landslide test runs as follows.

\begin{enumerate}
	\item Instrument the test program by inspecting its binary,
		learning the addresses of important functions, global variables, et cetera
		(\cref{sec:landslide-glue}).
	\item Execute the program under simulation.
		\begin{enumerate}
			\item At each instruction:
				\begin{enumerate}
					\item Update the {\em scheduler},
						a state machine depicting the runnable and/or blocked
						threads currently existing in the simulated program,
						as well as a set of action flags to track what each thread is up to,
						such as waiting to lock a mutex at a given memory address
						(\cref{sec:landslide-scheduler}).
					\item Record reads and writes to shared memory (\cref{sec:landslide-memory}).
					\item Check whether the current program state constitutes a bug,
						and if so,
						emit a preemption trace (\cref{sec:landslide-foundabug})
						and halt Landslide's execution.
						Bug-detection predicates range from
						simply checking if the program tripped an assert
						or accessed invalid memory,
						to checking the above-mentioned memory accesses against the current
						heap allocation state for use-after-frees (\cref{sec:landslide-valgrind-mode}),
						to querying the scheduler for deadlocks (\cref{sec:landslide-fp-deadlock}),
						to checking heuristically for infinite loops and livelock
						(\cref{sec:landslide-infloop}).
					\item Check whether the current program state is a preemption point
						(\cref{sec:landslide-pps}).
					\item Query the scheduler to detect when the test has completed successfully,
						i.e., all threads have finished executing normally
						(\cref{sec:landslide-scheduler-statemachine}).
				\end{enumerate}
			\item At each preemption point, identified as in 2(a)iv above:
				\begin{enumerate}
					\item Check the set of memory accesses since the previous preemption point,
						recorded as in 2(a)ii above,
						for conflicts with other threads
						(\cref{sec:landslide-shm}),
						and further for data races (\cref{sec:landslide-datarace}).
					\item Select which thread should run next (\cref{sec:landslide-arbiter}),
						Heuristically detecting yield-blocked threads
						may inform this decision (\cref{sec:landslide-blocking-yield}).
					\item Checkpoint the execution state in case future executions
						should wish to rewind and try a different thread from the one first chosen here
						(\cref{sec:landslide-save}, \cref{sec:landslide-timetravel}).
						% may be informed by heuristix/yield blocking
					\item Force the chosen thread to run by injecting timer interrupts
						(\cref{sec:landslide-interrupce},
						\cref{sec:landslide-scheduler-interrupce}).
				\end{enumerate}
				Note that Landslide maintains the invariant that each transition between two preemption points
				consist of instructions executed by exactly one thread;
				i.e., every thread switch must be punctuated by a preemption point.
				% preemption point one thread per invariant...
		\end{enumerate}
	\item At the end of each execution, identified as in 2(a)v above:
		\begin{enumerate}
			\item Analyze the set of memory conflicts,
				computed as in 2(b)i above,
				using Dynamic Partial Order Reduction (DPOR)
				to decide which preemption point to backtrack to and
				% once there
				%% once then % lol
				which new thread to run
				(\cref{sec:landslide-dpor}).
				This may be constrained by Iterative Context Bounding (\cref{sec:landslide-icb}).
			\item Estimate the structure of the resulting state space
				to predict overall runtime and total interleavings needed to test
				(\cref{sec:landslide-estimate}).
			\item If DPOR selected an alternate interleaving to explore,
				backtrack the simulation state to that preemption point
				(\cref{sec:landslide-timetravel})
				force the new thread to run
				(\cref{sec:landslide-interrupce}, \cref{sec:landslide-scheduler-interrupce}),
				and repeat from step 2.
				%If DPOR produced no results,
				Otherwise, halt Landslide's execution,
				declaring the explored state space free of bugs.
		\end{enumerate}
\end{enumerate}
}

As of this thesis's writing, Landslide supports the use of two possible simulators:

@@

\subsubsection{Disclaimer}

@@

\subsubsection{State machine}
\label{sec:landslide-scheduler-statemachine}

The first part of {\tt sched\_update()} is to update the state machine of thread actions and runnability.
Much of this functionality is found in
{\tt sched\_update\_kern\_state\_machine()}
and
{\tt sched\_update\_user\_state\_machine()}.
The current intruction pointer is compared against the known locations of
the mutex API, system calls, runnable/descheduling {\tt tell\_\allowbreak{}landslide} annotations, and so on,
and locksets, action flags, and runqueue membership are updated accordingly.
\revision{Landslide also queries the scheduler state after it updates every instruction,
via {\tt test\_update\_state()} ({\tt test.c}),
to check the existence and/or runnability of all the system's threads
and determine whether or not the test case has completed execution.}

\subsubsection{Interrupt injection}
\label{sec:landslide-scheduler-interrupce}

@@

\subsubsection{Memory conflicts}
\label{sec:landslide-shm}

@@

% time travel, explore, arbiter
\subsection{Bug-finding output}
\label{sec:landslide-foundabug}

@@

\subsubsection{Infinite loop detection}
\label{sec:landslide-infloop}

@@

\subsubsection{Yield-loop detection}
\label{sec:landslide-blocking-yield}

@@

\subsubsection{False-positive deadlock avoidance}
\label{sec:landslide-fp-deadlock}
